Visual Inference Lineups

### 2020-01-12 paper/paper.Rmd
rename bibs

---
title: "Comparing the Effectiveness of the Choropleth Map with a Hexagon Tile Map for Communicating Cancer Statistics"
affiliation:
  author-columnar: true         ## one column per author
  #institution-columnar: true  ## one column per institution (multiple autors eventually)
  # wide: true                  ## one column wide author/affiliation fields
  institution:
    - name: Queensland University of Technology
      department: Science and Engineering Faculty
      location: Brisbane, Australia
      email: stephanie.kobakian@qut.edu.au
      mark: 1
      author:
        - name: Stephanie Kobakian
    - name: Monash University
      department: Econometrics and Business Statistics Faculty
      location: Melbourne, Australia
      email: dicook@monash.edu
      mark: 1
      author:
        - name: Dianne Cook
keywords: ["statistics", "visual inference", "geospatial", "population"]
abstract: |
  The choropleth map display is commonly used for communicating spatial distributions across geographic areas. However, when choropleths are used, the sizes of areas can lead to misinterpretation of the distribution. The visualisation method used to present geospatial data will influence the understanding of the distribution derived by map users. Choosing an effective alternative display could positively influence the communication of the spatial distribution. The hexagon tile map is presented as an alternative display for effectively visualising population related distributions. Visual inference is used to measure the power of design, and the choropleth is used as a comparison. The hexagon tile map display is also tested using a distribution that is directly related to the geography, with values monotonically decreasing from the North West to South East areas of Australia. This study finds the single map in a hexagon tile map lineup that contains a population related distribution is detected with greater probability than the same data displayed in a choropleth map. These findings should encourage map creators to implement alternative displays, and consider a hexagon tile map when presenting spatial distributions of heterogeneous areas.
dev: png
bibliography: mybibfile.bib
output:
  bookdown::pdf_book: 
      base_format: rticles::ieee_article
      number_sections: true
editor_options: 
  chunk_output_type: inline
link-citations: true
citation_package: biblatex
csl: springer-vancouver.csl
header-includes: \usepackage{graphicx}
---

@@

The analysis of the work was completed in R [@RCore] with the use of the following packages:
 - For document creation: rmarkdown [@rmarkdown], rticles [@rticles], knitr [@knitr].
 - For lineup creation and data analysis: tidyverse [@tidyverse], nullabor [@nullabor], ggthemes [@ggthemes], RColorBrewer [@RColorBrewer].
 - For image displays: cowplot [@cowplot], png [@png], grid [@grid].
 - For modeling and presentation of models: lme4 [@lme4], kableExtra [@kableExtra].

>>>>

---
title: "Comparing the Effectiveness of the Choropleth Map with a Hexagon Tile Map for Communicating Cancer Statistics"
affiliation:
  author-columnar: true         ## one column per author
  #institution-columnar: true  ## one column per institution (multiple autors eventually)
  # wide: true                  ## one column wide author/affiliation fields
  institution:
    - name: Queensland University of Technology
      department: Science and Engineering Faculty
      location: Brisbane, Australia
      email: stephanie.kobakian@qut.edu.au
      mark: 1
      author:
        - name: Stephanie Kobakian
    - name: Monash University
      department: Econometrics and Business Statistics Faculty
      location: Melbourne, Australia
      email: dicook@monash.edu
      mark: 1
      author:
        - name: Dianne Cook
keywords: ["statistics", "visual inference", "geospatial", "population"]
abstract: |
  The choropleth map display is commonly used for communicating spatial distributions across geographic areas. However, when choropleths are used, the sizes of areas can lead to misinterpretation of the distribution. The visualisation method used to present geospatial data will influence the understanding of the distribution derived by map users. Choosing an effective alternative display could positively influence the communication of the spatial distribution. The hexagon tile map is presented as an alternative display for effectively visualising population related distributions. Visual inference is used to measure the power of design, and the choropleth is used as a comparison. The hexagon tile map display is also tested using a distribution that is directly related to the geography, with values monotonically decreasing from the North West to South East areas of Australia. This study finds the single map in a hexagon tile map lineup that contains a population related distribution is detected with greater probability than the same data displayed in a choropleth map. These findings should encourage map creators to implement alternative displays, and consider a hexagon tile map when presenting spatial distributions of heterogeneous areas.
dev: pdf
bibliography: paper.bib
output:
  bookdown::pdf_book: 
      base_format: rticles::ieee_article
      number_sections: true
editor_options: 
  chunk_output_type: inline
link-citations: true
citation_package: biblatex
csl: springer-vancouver.csl
header-includes: \usepackage{graphicx}
---

@@

The analysis of the work was completed in R [@RCore] with the use of the following packages:

 - For document creation: rmarkdown [@rmarkdown], rticles [@rticles], knitr [@knitr].
 - For lineup creation and data analysis: tidyverse [@tidyverse], nullabor [@nullabor], ggthemes [@ggthemes], RColorBrewer [@RColorBrewer].
 - For image displays: cowplot [@cowplot], png [@png], grid [@grid].
 - For modeling and presentation of models: lme4 [@lme4], kableExtra [@kableExtra].

### 2020-01-12 paper/supplementary/lineups.Rmd
rename bibs

---
title: "Visual Inference Lineups"
author: "Stephanie Kobakian and Dianne Cook"
output: bookdown::pdf_document
link_citatons: true
---

@@

The choices made by participants are examined in Fig. \ref{fig:choices}.
Participants were misled by the choropleth display, but not the hexagon display for All Cities displays except (2). The maps with a North West to South East trend was chosen with much greater frequency in all displays. All of Three Cities displays, except (4), were detected in the hexagon display. All except one lineup had at least one participant select the correct map in the lineup as shown in Fig. \ref{fig:choices}.

```{r choices, fig.cap = "Each facet is associated with one lineup, the height of the points show the proportion of the participants that made each choice when considering each lineup. The points coloured orange show the map which contained a trend model, these are the correct choices. The numbers differentiate the replicates of each trend model and type of map display. Participants were able to select 0 to indicate they did not want to choose a map.", fig.height = 8, fig.width =6}
d %>% 
  count(type, trend, choice, replicate, choice, detect_f) %>% 
  group_by(type, trend, replicate) %>% 
  mutate(prop = n/sum(n)) %>% 
  mutate(repl = paste("Rep:", replicate, ":\n", type,  sep = "")) %>%
  mutate(bottom = 0) %>% 
  ggplot() + 
  geom_point(aes(x = choice, y = prop, color = detect_f), size = 3) + 
  geom_segment(aes(x = choice, xend = choice,y = bottom, yend = prop, colour = detect_f), size = 1) +
  facet_grid(repl ~ trend, 
    drop = TRUE, as.table = TRUE, scales = "free_y") +
  labs(x = "Choice of plot in lineup", y = "Amount of choices") +
  scale_colour_manual(values = detect_colours) +
  scale_x_continuous(breaks = seq(from = 0, to = 12)) +
  scale_y_continuous(breaks = seq(from = 0.0, to = 1.0, by = 0.1)) +
  theme(legend.position = "bottom") + 
  guides(colour = FALSE, fill = FALSE) +
  theme(strip.text.y = element_text(angle = 0))
```

```{r}
choice_tabs <- d %>% 
  count(trend, type, choice, replicate, choice, detect) %>% 
  group_by(trend, type, replicate) %>% 
  mutate(prop = round(n/sum(n), 2)) %>% 
  mutate(repl = paste("Rep:", replicate, ":\n", type,  sep = "")) %>%
  mutate(bottom = 0) %>% 
  select(Trend = trend, Rep = replicate, Type = type, prop, choice) %>% 
  spread(choice, prop, fill = "0") %>% 
  nest(-Trend)
```

>>>>

---
title: "Visual Inference Lineups"
author: "Stephanie Kobakian and Dianne Cook"
output: pdf_document
bibliography: supplementary.bib
link-citations: true
citation_package: biblatex
csl: springer-vancouver.csl
---

@@

Each lineup had twelve map displays, this gave participants the choice of any plot, and the choice to not provide a response. this non-reponse is indicated by 0.
The choices made by participants are displayed in Fig. \ref{fig:choices}. The height of each orange lollipop indiates the proportion of participants that selected the map display of real data, they represent the correct choices.
The green lollipops show the proportion of participants that selected the incorrect displays in each lineup.

The proportion of choices are also presented separately for each trend model in Table. \ref{tab:choice-nwse}, Table. \ref{tab:choice-three}, and  Table. \ref{tab:choice-all}.
The correct map display in lineups with a North West to South East trend was chosen correctly with much greater frequency. 
In the lineups of All Cities displays, participants were misled by the choropleth display, but not the hexagon display for all except (2). 
All of Three Cities displays, except (4), were detected in the hexagon display. All except one lineup had at least one participant select the correct map in the lineup as shown in Fig. \ref{fig:choices}.

```{r choices, fig.cap = "Each facet is associated with one lineup, the height of the points show the proportion of the participants that made each choice when considering each lineup. The points coloured orange show the map which contained a trend model, these are the correct choices. The numbers differentiate the replicates of each trend model and type of map display. Participants were able to select 0 to indicate they did not want to choose a map.", fig.height = 8, fig.width =6, results = "asis"}
d %>% 
  count(type, trend, choice, replicate, choice, detect_f) %>% 
  group_by(type, trend, replicate) %>% 
  mutate(prop = n/sum(n)) %>% 
  mutate(repl = paste("Rep:", replicate, ":\n", type,  sep = "")) %>%
  mutate(bottom = 0) %>% 
  ggplot() + 
  geom_point(aes(x = choice, y = prop, color = detect_f), size = 3) + 
  geom_segment(aes(x = choice, xend = choice,y = bottom, yend = prop, colour = detect_f), size = 1) +
  facet_grid(repl ~ trend, 
    drop = TRUE, as.table = TRUE, scales = "free_y") +
  labs(x = "Choice of plot in lineup", y = "Amount of choices") +
  scale_colour_manual(values = detect_colours) +
  scale_x_continuous(breaks = seq(from = 0, to = 12)) +
  scale_y_continuous(breaks = seq(from = 0.0, to = 1.0, by = 0.1)) +
  theme(legend.position = "bottom") + 
  guides(colour = FALSE, fill = FALSE) +
  theme(strip.text.y = element_text(angle = 0))
```

```{r}
choice_tabs <- d %>% 
  count(trend, type, choice, replicate, choice, detect) %>% 
  group_by(trend, type, replicate) %>% 
  mutate(prop = round(n/sum(n), 2)) %>% 
  mutate(repl = paste("Rep:", replicate, ":\n", type,  sep = "")) %>%
  mutate(bottom = 0) %>% 
  select(Trend = trend, Rep = replicate, Type = type, prop, choice) %>% 
  spread(choice, prop, fill = "0") %>% 
  nest(data = c(Rep, Type, `0`, `1`, `2`, `3`, `4`, `5`, `6`, `7`, `8`, `9`, 
    `10`, `11`, `12`))
```
